---
title: Forecasting is not prophecy
author: Dealing with high-dimensional probabilistic forecasts in&nbsp;practice
date: Rob J Hyndman
classoption: compress
titlefontsize: 28pt
toc: false
output:
  binb::monash:
    incremental: no
    fig_height: 4.8
    keep_tex: no
    includes:
      in_header: preamble.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, cache=TRUE, dev.args=list(bg=grey(0.9), pointsize=11))
options(width=67)
set.seed(20190714)
library(fpp3)
aus_holidays <- tourism %>%
  mutate(
    State = recode(State,
                   "Australian Capital Territory" = "ACT",
                   "New South Wales"="NSW",
                   "Northern Territory" = "NT",
                   "Queensland" = "QLD",
                   "South Australia" = "SA",
                   "Tasmania" = "TAS",
                   "Victoria"="VIC",
                   "Western Australia" = "WA"
    )
  ) %>%
  filter(Purpose == "Holiday") %>% select(-Purpose)
austa <- aus_holidays %>%
  summarize(Trips = sum(Trips))
```

## What is it?

\placefig{5.1}{1.5}{width=7.7cm}{sheepliverbm}

\only<2>{\begin{textblock}{4.5}(0.5,1.7) \raggedright\fontsize{17}{19}\sf
\textbf{Clay model of sheep's liver}\\[0.4cm]
Used by Bab\-ylon\-ian forecasters approximately 600~B.C.\\[.9cm]
Now in \rlap{British Museum.}
\end{textblock}}

## Modern prophecy

\placefig{0}{1.3}{width=12.9cm, height=20cm}{bad_forecasts}

## Probabilistic forecasting

```{r austa0, fig.width=7.88, fig.height=6}
ggplot(austa, aes(x = Quarter)) +
  geom_line(aes(y = Trips, colour = "Data"), colour='black') +
  geom_point(
    data=tsibble(Quarter=yearquarter("2021 Jan"),Trips=15190, index=Quarter),
    aes(y=Trips,x=Quarter),
    alpha=0
  ) +
  ylab("Thousands of trips") + xlab("Quarter") +
  ggtitle("Total domestic holiday trips: Australia") +
  guides(colour=FALSE)
```

## Probabilistic forecasting

```{r austa1}
fit <- austa %>% model(ETS(Trips))
```

```{r austa1a, fig.width=9, fig.height=6, dependson="austa1"}
sim <- fit %>% generate(h = 10, times = 10) %>%
  mutate(
    replicate = factor(.rep, levels = 1:10, labels = paste("Future", 1:10))
  )
ggplot(austa, aes(x = Quarter)) +
  geom_line(aes(y = Trips, colour = "Data")) +
  geom_line(aes(y = .sim, colour = replicate), data = sim) +
  geom_point(
    data=tsibble(Quarter=yearquarter("2021 Jan"),Trips=15190, index=Quarter),
    aes(y=Trips,x=Quarter),
    alpha=0
  ) +
  ylab("Thousands of trips") + xlab("Quarter") +
  ggtitle("Total domestic holiday trips: Australia") +
  scale_colour_manual(values = c("#000000", rainbow(10)),
                      breaks = c("Data", paste("Future", 1:10)),
                      name = " ")
```

## Probabilistic forecasting

```{r austa2, fig.width=8.63, fig.height=6, dependson="austa1"}
fit %>%
  forecast(h = 10) %>%
  autoplot(austa) +
  geom_point(
    data=tsibble(Quarter=yearquarter("2021 Jan"),Trips=15190, index=Quarter),
    aes(y=Trips,x=Quarter),
    alpha=0
  ) +
  ylab("Thousands of trips") + xlab("Quarter") +
  ggtitle("Total domestic holiday trips: Australia")
```

## Multivariate probabilistic forecasting

```{r austa3, fig.height=8, fig.width=6, out.height="92%"}
z <- aus_holidays %>%
  group_by(State) %>%
  summarize(Trips = sum(Trips)) %>%
  aggregate_key(State, Trips = sum(Trips))
z %>%
  model(ets = ETS(Trips)) %>%
  forecast(h = "3 years") %>%
  autoplot(z) +
  ylab("Thousands of trips")
```

\only<2>{\begin{textblock}{4.5}(7.5,3)\fontsize{11}{14}\sf
\begin{block}{}
These show the marginal distributions of the 9-dimensional distribution. In practice, we also want forecast for smaller geographic areas as well.
\end{block}
\end{textblock}}

## \fontsize{17}{16}\sf\bfseries Who needs multivariate probabilistic forecasts?

 * Tourism authorities forecasting visitor numbers to plan facilities.
 * Manufacturing companies forecasting product demand to plan their supply chains.
 * Call centres forecasting call volume to inform staff scheduling.
 * Technology companies forecasting web traffic to maintain service levels.
 * Energy companies forecasting electricity demand to prevent blackouts.

## \fontsize{16}{16}\sf\bfseries Challenges in multivariate probabilistic forecasting
\fontsize{12}{14}\sf

 1. How to produce high-dim probabilistic forecasts?
 2. How to use and interpret high-dim probabilistic forecasts?
 3. How to measure accuracy of high-dim probabilistic forecasts?

\pause

###
 * Users care about marginal distributions, but we need multivariate distribution to compute them.
 * Forecast reconciliation provides a way to efficiently compute marginal distributions.
 * Probability scoring allows us to measure distributional forecast accuracy.

## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

```{r score1, fig.height=3.9}
y <- 4.8
s <- function(y,i,pi)
{
  (1-i/100)*(pi-y) * (y < pi) + i/100*(y-pi) * (y >= pi)
}
x <- seq(-7,7,l=10001)
dx <- dnorm(x,3,1)
par(mar=c(3,2,1,5))
plot(x,3*dx, xlim=c(0,6), ylim=c(0,2), type="n",
     yaxt="n", yaxs='i', ylab="", xlab="",
     main="Forecast distribution")
polygon(c(x,x[1]),c(dx,dx[1])*3,border=FALSE,col='gray')
axis(1,at=y,labels="y",line=1,tck=0.06, lwd=2, col='blue')
```

## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

```{r score2, fig.height=3.9}
p <- 0.2
poe <- qnorm(p,3,1)
x2 <- x[x>=poe]
dx2 <- dx[x>=poe]
par(mar=c(3,2,1,5))
plot(x,3*dx, xlim=c(0,6), ylim=c(0,2), type="n",
     yaxt="n", yaxs='i', ylab="", xlab="",
     main="Forecast distribution")
polygon(c(x,x[1]),c(dx,dx[1])*3,border=FALSE,col='gray')
polygon(c(x2[1],x2,x2[1]),3*c(0,dx2,0),col='red',border=TRUE)
lines(x,s(x,p*100, poe),lwd=2,col='DarkGreen')
text(1.6,.1,"20%",col='white')
text(3.6,.2,"80%",col='black')
axis(1,at=poe,labels=expression(q[0.2]),line=1,tck=0.2)
axis(1,at=y,labels="y",line=1,tck=0.06, lwd=2, col='blue')
```

## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

```{r score2a, fig.height=3.9}
p <- 0.2
poe <- qnorm(p,3,1)
x2 <- x[x>=poe]
dx2 <- dx[x>=poe]
par(mar=c(3,2,1,5))
plot(x,3*dx, xlim=c(0,6), ylim=c(0,2), type="n",
     yaxt="n", yaxs='i', ylab="", xlab="",
     main="Forecast distribution")
polygon(c(x,x[1]),c(dx,dx[1])*3,border=FALSE,col='gray')
polygon(c(x2[1],x2,x2[1]),3*c(0,dx2,0),col='red',border=TRUE)
lines(x,s(x,p*100, poe),lwd=2,col='DarkGreen')
text(1.6,.1,"20%",col='white')
text(3.6,.2,"80%",col='black')
axis(1,at=poe,labels=expression(q[0.2]),line=1,tck=0.2)
axis(1,at=y,labels="y",line=1,tck=0.2, col='blue')
lines(rep(y,2), c(0,s(y,100*p,poe)), col='blue')
lines(c(y,10), rep(s(y,100*p,poe),2), col='blue')
axis(4,at=s(y,100*p,poe), labels=expression(s[0.2](q[0.2],y)),las=1)
```


## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

```{r score3, fig.height=3.9}
poe <- qnorm(0.4,3,1)
x2 <- x[x>=poe]
dx2 <- dx[x>=poe]
par(mar=c(3,2,1,5))
plot(x,3*dx, xlim=c(0,6), ylim=c(0,2), type="n",
     yaxt="n", yaxs='i', ylab="", xlab="",
     main="Forecast distribution")
polygon(c(x,x[1]),c(dx,dx[1])*3,border=FALSE,col='gray')
polygon(c(x2[1],x2,x2[1]),3*c(0,dx2,0),col='red',border=TRUE)
lines(x,s(x,40, poe),lwd=2,col='DarkGreen')
text(1.8,.1,"40%",col='white')
text(3.6,.2,"60%",col='black')
axis(1,at=poe,labels=expression(q[0.4]),line=1,tck=0.2)
axis(1,at=y,labels="y",line=1,tck=0.2, col='blue')
lines(rep(y,2), c(0,s(y,40,poe)), col='blue')
lines(c(y,10), rep(s(y,40,poe),2), col='blue')
axis(4,at=s(y,40,poe), labels=expression(s[0.4](q[0.4],y)),las=1)
```


## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

```{r score4, fig.height=3.9}
p <- 0.9
poe <- qnorm(p,3,1)
x2 <- x[x>=poe]
dx2 <- dx[x>=poe]
par(mar=c(3,2,1,5))
plot(x,3*dx, xlim=c(0,6), ylim=c(0,2), type="n",
     yaxt="n", yaxs='i', ylab="", xlab="",
     main="Forecast distribution")
polygon(c(x,x[1]),c(dx,dx[1])*3,border=FALSE,col='gray')
polygon(c(x2[1],x2,x2[1]),3*c(0,dx2,0),col='red',border=TRUE)
lines(x,s(x,p*100, poe),lwd=2,col='DarkGreen')
text(4.6,.1,"10%",col='white')
text(3.6,.2,"90%",col='black')
axis(1,at=poe,labels=expression(q[0.9]),line=1,tck=0.2)
axis(1,at=y,labels="y",line=1,tck=0.2, col='blue')
lines(rep(y,2), c(0,s(y,p*100,poe)), col='blue')
lines(c(y,10), rep(s(y,p*100,poe),2), col='blue')
axis(4,at=s(y,p*100,poe), labels=expression(s[0.9](q[0.9],y)),las=1)
```


## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.


\begin{block}{Quantile Score (pinball loss)}\vspace*{-0.1cm}
$$
s_p(q_p,y) = \begin{cases}
(1 - p) (q_p - y), & \text{if $y < q_p$};\\
p (y - q_p), & \text{if $y\ge q_p$}.
\end{cases}
$$
\end{block}

\pause

\begin{block}{Continuous Rank Probability Score}
$$
\text{CRPS}(F,y) = 2 \int_0^1 s_p(q_p,y) dp.
$$
\end{block}


## Probability scoring

 * $y=$ actual value
 * $F=$ forecast distribution
 * $q_p = F^{-1}(p) =$ quantile with probability $p$.

\begin{block}{Continuous ranked probability scores}\vspace*{-.5cm}
\begin{align*}
  \text{CRPS}(F,y) & = 2 \int_0^1 s_p(q_p,y) dp.\\
   &= \int_{-\infty}^\infty (F(x) -1\{y \le x\})^2 dx
\end{align*}
\end{block}\vspace*{-0.3cm}\pause

 * CRPS in same units as observations
 * Expected CRPS minimized when observation $y$ from same distribution as $F$.
 * Maximizes sharpness subject to calibration.


## Tidyverts R packages

\begin{textblock}{3.8}(8,0)\begin{alertblock}{}\Large\textbf{tidyverts.org}\end{alertblock}\end{textblock}

\placefig{1}{1.4}{width=4cm}{tsibble.png}
\placefig{5}{1.4}{width=4cm}{tsibbledata.png}
\placefig{3}{4.85}{width=4cm}{feasts.png}
\placefig{7}{4.85}{width=4cm}{fable.png}

## `tsibble` objects

\fontsize{9.5}{11}\sf

```{r tsibble, echo = TRUE}
library(tsibble)
library(tsibbledata)
aus_holidays
```

\only<2->{\begin{textblock}{.97}(2.05,4.17)
\begin{alertblock}{}\fontsize{9}{9}\sf Index\phantom{dg}\end{alertblock}
\end{textblock}}
\only<3->{\begin{textblock}{2.2}(3.42,4.17)
\begin{alertblock}{}\fontsize{9}{9}\sf Keys\phantom{dg}\end{alertblock}
\end{textblock}}
\only<4-5>{\begin{textblock}{1.5}(5.99,4.17)
\begin{alertblock}{}\fontsize{9}{9}\sf Measure\phantom{dg}\end{alertblock}
\end{textblock}}

\only<5>{\begin{textblock}{3}(9,5)
\begin{block}{}\fontsize{10}{10}\sf Domestic overnight holiday trips in thousands by state/region.\phantom{dg}\end{block}
\end{textblock}}

## High-dimensional time series data
\fontsize{9.5}{11}\sf

```{r tourism, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) # 2 year test set
```

## Compute all aggregates
\fontsize{9.5}{11}\sf

```{r aggregates, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips))
```

## Fit univariate models
\fontsize{9.5}{11}\sf

```{r models, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips))
```

## Reconcile results
\fontsize{9.5}{11}\sf

```{r coherent, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(ets = min_trace(ets))
```

## Produce probabilistic forecasts
\fontsize{9.5}{11}\sf

```{r forecasts, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(ets = min_trace(ets)) %>%
  forecast(h = "2 years")
```

```{r crps, results='hide'}
get_sd <- function(x) {
  purrr::map_dbl(x, function(x){x$sd})
}
Gaussian_crps <- function(y, mean, sd) {
  z <- (y-mean)/sd
  sd*(z*(2*pnorm(z)-1)+2*dnorm(z)-1/sqrt(pi))
}
fc <- aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(ets = min_trace(ets)) %>%
  forecast(h = "2 years")
y <- filter(aus_holidays, Quarter >= yearquarter("2016 Q1")) %>%
  aggregate_key(State/Region, Trips=sum(Trips))
ac <- fc %>%
  mutate(sd = get_sd(.distribution)) %>%
  as_tibble() %>%
  rename(Mean = Trips) %>%
  left_join(y) %>%
  mutate(
    z = (Trips - Mean)/sd,
    CRPS = Gaussian_crps(Trips, Mean, sd)
  ) %>%
  group_by(State, Region, .model) %>%
  summarize(CRPS = mean(CRPS)) %>%
  select(.model, Region, State, everything())
ac1 <- accuracy(fc, y) %>%
  left_join(ac) %>%
  select(.model, Region, State, CRPS) #%>%
  #filter(!is_aggregated(State) & !is_aggregated(Region))
ac2 <- ac1 %>%
  spread(key=.model, value=CRPS)
ac3 <- ac2 %>%
  mutate(RelCRPS = ets/mean)
ac4 <- ac3 %>%
  arrange(RelCRPS)
```

## Compare against test set
\fontsize{9.5}{11}\sf

```{r accuracy, eval=FALSE, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(coherent = min_trace(ets)) %>%
  forecast(h = "2 years") %>%
  accuracy(aus_holidays)
```

```{r accuracy2, eval=TRUE, echo=FALSE, dependson='crps'}
ac1
```

## Probability scoring
\fontsize{9.5}{11}\sf

```{r crps2, eval=FALSE, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(coherent = min_trace(ets)) %>%
  forecast(h = "2 years") %>%
  accuracy(aus_holidays) %>%
  spread(key=.model, value=CRPS)
```

```{r crps2a, eval=TRUE, echo=FALSE, dependson='crps'}
ac2
```

## Probability scoring
\fontsize{9.5}{11}\sf

```{r crps3, eval=FALSE, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(coherent = min_trace(ets)) %>%
  forecast(h = "2 years") %>%
  accuracy(aus_holidays) %>%
  spread(key=.model, value=CRPS) %>%
  mutate(RelCRPS = coherent/mean)
```

```{r crps3a, eval=TRUE, echo=FALSE, dependson='crps'}
ac3
```

## Probability scoring
\fontsize{9.5}{11}\sf

```{r crps4, eval=FALSE, echo=TRUE}
aus_holidays %>%
  filter(Quarter <= yearquarter("2015 Q4")) %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  model(mean = MEAN(Trips), ets = ETS(Trips)) %>%
  reconcile(coherent = min_trace(ets)) %>%
  forecast(h = "2 years") %>%
  accuracy(aus_holidays) %>%
  spread(key=.model, value=CRPS) %>%
  mutate(RelCRPS = coherent/mean) %>%
  arrange(RelCRPS)
```

```{r crps4a, eval=TRUE, echo=FALSE, dependson='crps'}
ac4
crps <- ac4
```

## Best forecast (RelCRPS)

```{r best, dependson='crps'}
best <- crps %>%
  filter(RelCRPS==min(RelCRPS)) %>%
  select(Region, State)
data <- aus_holidays %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  right_join(best)
region <- as.character(best$Region)
region <- if_else(is.na(region), "<total>", region)
fcast <- fc %>%
  filter(.model=='ets') %>%
  right_join(best) %>%
  as_fable(resp=Trips, dist=.distribution)
autoplot(fcast, colour='red') +
  autolayer(data) +
  ggtitle(paste0("Domestic trips: ", best$State, "/", region))
```

## Worst forecast (RelCRPS)

```{r worst, dependson='crps'}
worst <- crps %>%
  filter(RelCRPS==max(RelCRPS)) %>%
  select(Region, State)
region <- as.character(worst$Region)
region <- if_else(is.na(region), "<total>", region)
data <- aus_holidays %>%
  aggregate_key(State/Region, Trips=sum(Trips)) %>%
  right_join(worst)
fcast <- fc %>%
  filter(.model=='ets') %>%
  right_join(worst) %>%
  as_fable(resp=Trips, dist=.distribution)
autoplot(fcast, colour='red') +
  autolayer(data) +
  ggtitle(paste0("Domestic trips: ", worst$State, "/", region))
```


## \fontsize{16}{16}\sf\bfseries Challenges in multivariate probabilistic forecasting
\fontsize{12}{14}\sf

 1. The fable package produces high-dim probabilistic forecasts?
 2. We can focus on marginal distributions for interpretation, while needing the full distribution for calculation.
 3. Forecast reconciliation provides a way to efficiently compute marginal distributions.
 4. Probability scoring allows us to measure distributional forecast accuracy.

## Acknowledgements


\begin{block}{}
\fontsize{11}{11}\sf
\centering\begin{tabular}{l@{\hspace*{1cm}}l}
\includegraphics[height=4cm, width=10cm]{mitch2} &
\includegraphics[height=4cm, width=10cm]{earowang} \\
Mitchell O'Hara-Wild &
Earo Wang \\
\end{tabular}
\end{block}\vspace*{0.5cm}

\begin{alertblock}{}
\centerline{\textbf{tidyverts.org}}
\centerline{\textbf{robjhyndman.com}}
\end{alertblock}
